---
title: "Functions-R"
output: html_notebook
---
This notebook serves the purpose of holding the functions that have been created within the
class. Each function will have a descriptive name such that is may be understood
out of context. 

# Functions from r-Intro

The function below computes the number of handshakes, h, that would occur with
a certain number of people, n. It then plots the number of people and connects
them via line segments.
```{r}
handShake = function(n=1, plotMe =FALSE)
{
  if(n < 1) {stop("n must have a value of at least 1");}
  if(!is.wholenumber(n)) {stop("n must be an integer");}
  h = n*(n-1)/2;
  if(plotMe) 
     {
     r = n
     plot(0,0,xlim=c(-n-1,n+1), ylim=c(-n-1,n+1), main=paste0(h," handshakes for ",n," people"), xlab="", ylab="", xaxt="n", yaxt="n",bty="n", asp=1);
     draw.circle(0,0,n,n+100);
     #draw.circle(0,0,n,n);
     #points(0,r, pch=16)
     #plot(r*sin(theta),r*cos(theta));
     #handshake.points = data.matrix(a = 1:n, b = )
     for(i in 1:n){
        deg = 360/n;
        theta = 2*pi*deg/360;
        plot(r*sin(i*theta),r*cos(i*theta), pch=16, )
        lines(r*sin(i*theta), r*cos(i*theta), type="l")
     }
  }
}
```

This function removes white space in a txt file, changes everything to lowercase
and then counts the number of each letter (a-z). It also counts the number of 
characters that do not count as a letter and puts them into "other". Every time
it counts a letter, it subtracts it from the file and moves to the next.
```{r}

countLetterInString = function(str, letter)
{
  nchar(as.character(str)) -nchar( gsub(letter, "", str, fixed=TRUE))
}
AlphabetCounter = function(str)
{
  str = gsub("[[:space:]]", "", str)
  str = tolower(str)
  df = data.frame(matrix(0, nrow=1, ncol=27, byrow=TRUE))
  colnames(df) = c(letters,"OTHER")
  for(letter in letters)
  {
    idx = which(letters == letter)
    df[1,idx] = countLetterInString(str,letter)
    str = gsub(letter,"",str,fixed=TRUE)
  }
  
  df[1,27] = nchar(str)
  df;
}
```

This function computes the determinant of a 3 by 3 matrix.
```{r}
ComputeDeterminate3=function(myMatrix)
{
  nrow=nrow(myMatrix);
  ncol=ncol(myMatrix);
  if (nrow != ncol)
  {stop("The matrix must be square to compute the determinant");}
  if (nrow != 3)
  {stop("The matrix must be 3 by 3");}
  a=myMatrix[1,1];
  b=myMatrix[1,2];
  c=myMatrix[1,3];
  d=myMatrix[2,1];
  e=myMatrix[2,2];
  f=myMatrix[2,3];
  g=myMatrix[3,1];
  h=myMatrix[3,2];
  i=myMatrix[3,3];
  
  # https://crypto.interactive-maths.com/hill-cipher.html
  (a*(e*i-f*h))-(b*(d*i-f*g))+(c*(d*h-e*g))
}
```

